\simname 2.0 User's Guide

### 2009-04-09 doc/manual.tex
add Id tags
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@848 ce7aba0d-9f81-40a6-a495-cecae84b0515
remove bogus svn:executable properties
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@938 ce7aba0d-9f81-40a6-a495-cecae84b0515
edited
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@1202 ce7aba0d-9f81-40a6-a495-cecae84b0515

@@

The latest version of the simulator is available from
\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
users can extract this archive using the tar utility
\begin{verbatim}

  tar xvfz booksim-2.0.tar.gz  
\end{verbatim}

Windows users can use a compression program such as Winrar to extract
the archive.

The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
LEX and YACC tool (also known as FLEX and BISON) are needed to create
the configuration parser.  These are standard tools in any UNIX/Linux
development environment.  It is suggested that Windows users download
the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
development tools to simplify their compilation process.  The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools.  At Stanford, for example, the compiler, YACC, and
LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
\texttt{Makefile} reflects this:
\begin{verbatim}
CPP    = /usr/pubsw/bin/g++
YACC   = /usr/pubsw/bin/bison -d
LEX    = /usr/pubsw/bin/flex
\end{verbatim}
Then, the simulator can be compiled by running \texttt{make} in the
directory that contains the \texttt{Makefile}.

\subsection{Running a simulation}
\label{sec:run_example}

The syntax of the simulator is simply
\begin{verbatim}
  ./booksim [configfile]
\end{verbatim}
The optional parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.

@@

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and and four
virtual channels.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject 0.15 packets per simulation
cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}.

@@

\begin{figure}
\begin{verbatim}
% Average latency = 6.02008
% Accepted packets = 0.11 at node 52 (avg = 0.147094)
% latency change    = 1
% throughput change = 1
>>>>>>> variant B
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
Unable to open flit watch file, continuing with simulation
0%=================================
% Average latency = 10.6252
% Accepted packets = 0.119 at node 5 (avg = 0.148875)
lat(1) = 10.6252;
bins = [ 0 ...999];
freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]

...

@@

...

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6031 cycles
====== Traffic class 0 ======
Overall average latency = 10.658 (1 samples)
Overall average accepted rate = 0.148383 (1 samples)
Overall min accepted rate = 0.134279 (1 samples)
Average hops = 5.01172 (57148 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

@@

\section{Examples}
\label{sec:examples}

One of the most basic performance measures of any interconnection
network is its latency versus offered load.
Figure~\ref{fig:lat_vs_load} shows a simple configuration file for
making this measurement in a 8-ary 2-mesh network under the transpose
traffic pattern.  This configuration was used to generate Figure 25.2
in PPIN.  The particular configuration accounts for some small delays
and pipelining of the input-queued router and also introduces a small
input speedup to account for any inefficiencies in allocation.  By
running simulations for many increments of \texttt{injection\_rate},
the average latency curve can be found.  Then, to compare the
performance of dimension-order routing against several other routing
algorithms, for example, the \texttt{routing\_function} option can be
changed.

@@

sim_type       = latency;
injection_rate = 0.005;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file. Most configuration parameters have default values that should work without user specification

@@

\item[flattened butterfly] A topology based on ***

@@

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.

@@

\appendix

\section{Random number generation}

>>>>

% $Id$

% Copyright (c) 2007, Trustees of Leland Stanford Junior University
% All rights reserved.
% 
% Redistribution and use in source and binary forms, with or without 
% modification, are permitted provided that the following conditions are met:
% 
% Redistributions of source code must retain the above copyright notice, this 
% list of conditions and the following disclaimer.
% Redistributions in binary form must reproduce the above copyright notice, 
% this list of conditions and the following disclaimer in the documentation 
% and/or other materials provided with the distribution.
% Neither the name of the Stanford University nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
% 
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" 
% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE 
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE 
% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE 
% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR 
% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF 
% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS 
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN 
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) 
% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE 
% POSSIBILITY OF SUCH DAMAGE.

@@

The latest version of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run 
\begin{verbatim}
svn co https://booksim.svn.sourceforge.net/svnroot/booksim booksim 
\end{verbatim}

The simulator files should now be in \texttt{booksim/trunk/src/}. The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$). The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file. But if you do not plan on making changes to the front end parser, the LEX and YACC are not needed. The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools. The default \texttt{Makefile} should work on the Stanford Leland machines. Type \texttt{make} to build the simulator.

For Windows users: the above instructions has been tested to work with the newest version of Cygwin. 
\subsection{Running a simulation}
\label{sec:run_example}

The syntax of the simulator is simply
\begin{verbatim}
  ./booksim [configfile]
\end{verbatim}
The parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used. This particular configuration is stored in \texttt{examples/torus88}.

@@

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and four
virtual channels.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject (on average) 0.15 packets per simulation
cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}. Any parameters not specified by the user will take on a default value. The default value for every parameter in the simulator is in the file \texttt{booksim\_config.cpp}.

@@

\begin{figure}
\begin{verbatim}
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
0%=================================
% Average latency = 10.3206
% Accepted packets = 0.113 at node 38 (avg = 0.149031)
lat(1) = 10.3206;
bins = [ 0 ...999];
freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
 49 10 7 1 0 ....

....

@@

....

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6030 cycles
====== Traffic class 0 ======
Overall average latency = 10.2897 (1 samples)
Overall average accepted rate = 0.14834 (1 samples)
Overall min accepted rate = 0.129043 (1 samples)
Average hops = 4.99839 (57153 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

@@

\section{Example}
\label{sec:examples}

One of the most basic performance measures of any interconnection
network is its latency versus offered load.
Figure~\ref{fig:lat_vs_load} shows a simple configuration file for
making this measurement in a 8-ary 2-mesh network under the transpose
traffic pattern.  This configuration was used to generate a single data point in Figure 25.2
in PPIN.  The particular configuration accounts for some small delays
and pipelining of the input-queued router and also introduces a small
input speedup to account for any inefficiencies in allocation.  By
running simulations for many increments of \texttt{injection\_rate},
the average latency curve can be found.  Then, to compare the
performance of dimension-order routing against several other routing
algorithms, for example, the \texttt{routing\_function} option can be
changed.

@@

sim_type       = latency;
injection_rate = 0.01;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

All information used to configure a simulation is passed through a
configuration file as illustrated by the example in
Section~\ref{sec:run_example}.  This section lists the existing
configuration parameters --- a user can incorporate additional options
by changing the \texttt{booksim\_config.cpp} file.

@@

\item[flattened butterfly] A topology based on the paper "Flattened butterfly: a cost-efficient topology for high-radix networks" ISCA 2007

@@

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.

@@

\appendix
\section{Random number generation}
